Switch equivalence and weak prime domains for left-linear DPO rewriting systems

### 2025-03-07 full_version/permutations.tex
minori

	A \emph{left-linear DPO-rewriting system} is a pair $(\X, \R)$ where $\X$ is a $\mathcal{M}$-adhesive category and $R$ is a set of left $\mathcal{M}$-linear rules. $(\X, \R)$ will be called \emph{linear} if every rule in $R$ is $\mathcal{M}$-linear. Similarly, a \emph{consuming} left-linear DPO-rewriting system is one in which every rule is consuming.

	Given  two objects $G$ and $H$ and a rule $\rho=(l,r)$ in $\R$, a \emph{direct derivation $\mathscr{D}$ from $G$ to $H$ applying the rule $\rho$}, is a diagram as the one below, in which both squares are pushouts. 
	\[\xymatrix{L \ar[d]_{n}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H}\]
	The arrow $n$ is called the \emph{match} of the derivation, while $h$ is its \emph{back-match}.
	We will denote a direct derivation $\dder{D}$ between $G$ and $H$ as $\dder{D}\colon G\Mapsto H$. 
\end{definition}

@@

\begin{definition}
	Let $(\X, \R)$ be a left-linear DPO-rewrityng system, with $\X$ $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of $(\X, \R)$ is the (large)  directed graph $\gpo$ having as vertices the objects of $\X$ and in which an edge between $G$ and $H$ is a direct derivation $\dder{D}\colon G\Mapsto H$.	A \emph{derivation} $\der{D}$ between two objects $G$ and $H$ is a path between them in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are, respectively, $G$ and $H$.
\end{definition}

@@

	Moreover,  if every $\dder{D}_i$ applies the rule $\rho_i\in R$, then we can define an associated sequence of rules as $r(\der{D})$ as $\{\rho_i\}_{i=0}^n$.
\end{remark}

@@

\begin{example}\label{ex:1}
	Let us consider the category of $\cat{Graph}$ of all graphs and graph morphisms, which is adhesive (see, for instance \cite{lack2005adhesive,lack2006toposes}).   We can endow $\cat{Graph}$ with a set of left-linear rules, for instance taking the following ones.  Here numbers are used to represent the
	morphisms from the interface to the left- and right-hand
	sides. Rules $\rho_0$ and $\rho_1$ are linear: $\rho_0$ generates a
	new node and a new edge, while $\rho_1$ creates a self-loop edge. Instead,
	$\rho_2$ is left-linear but not linear, as it ``merges'' two nodes.
	\begin{center}
		\begin{tikzpicture}[node distance=2mm, font=\small]
			%, baseline=(current bounding box.center), font=\small]
			\node (l) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {};
					\node at (0,0) [node, label=below:$1$] (1) {} ;        
					%
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (l.north) {$L_0$};
			\node [right=of l] (k) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {};
					%
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (k.north) {$K_0$};
			\node[below] at (k.south) {$\rho_0$};
			\node  [right=of k] (r) {
				\begin{tikzpicture}
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {};
					\node at (.5,0) [node, label=below:$2$] (2) {};
					\draw[->] (1) to[out=20, in=160] (2);
					%
					\pgfBox
				\end{tikzpicture}
			};
			\node[font=\scriptsize, above] at (r.north) {$R_0$};
			\path (k) edge[->] node[trans, above] {} (l);
			\path (k) edge[->] node[trans, above] {} (r);    
		\end{tikzpicture}
		%  }
	%
	\hspace{1cm}
	%  \hfill
	%
	%  \subcaptionbox*{}{
		% RULE P1
		%
		\begin{tikzpicture}[node distance=2mm, font=\small]
			\node (l) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {} ;
					% 
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (l.north) {$L_1$};
			\node [right=of l] (k) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {};
					% 
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (k.north) {$K_1$};
			\node[below] at (k.south) {$\rho_1$};
			\node  [right=of k] (r) {
				\begin{tikzpicture}
					\node at (0,0) [node, label=below:$1$] (1) {}
					edge [in=55, out=85, loop] ();
					%
					\pgfBox
				\end{tikzpicture}
			};
			\node[font=\scriptsize, above] at (r.north) {$R_1$};
			\path (k) edge[->] node[trans, above] {} (l);
			\path (k) edge[->] node[trans, above] {} (r);
		\end{tikzpicture}
		%  }
	% 
	\hspace{1cm}
	%\hfill
	%
	%    \subcaptionbox*{}{
		% RULE P2
		%
		\begin{tikzpicture}[node distance=2mm, font=\small]
			\node (l) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {} ;
					\node at (0.5,0) [node, label=below:$2$] (2) {} ;
					%
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (l.north) {$L_2$};
			\node [right=of l] (k) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {} ;
					\node at (0.5,0) [node, label=below:$2$] (2) {} ;
					%
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (k.north) {$K_2$};
			\node[below] at (k.south) {$\rho_2$};
			\node  [right=of k] (r) {
				\begin{tikzpicture}
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$12$] (12) {};
					%
					\pgfBox
				\end{tikzpicture}
			};
			\node[font=\scriptsize, above] at (r.north) {$R_2$};
			\path (k) edge[->] node[trans, above] {} (l);
			\path (k) edge[->] node[trans, above] {} (r);
		\end{tikzpicture}
	\end{center}

@@

\begin{remark}
	By definition the inclusion $\ske \to \X$ is an equivalence. In particular, this mean that, for every objects $X$ of $\X$ there exists $\pi(X)$ in $\ske$ and an isomorphism $\phi_X\colon \pi(X) \to X$.
\end{remark}

\begin{proposition}\label{prop:ske}
	Every category $\X$ has a skeleton. 
\end{proposition}
\begin{proof}
	For every object $X\in \X$, pick a single representative $\pi(X)$ of its isomorphism class. Let $\ske$ be the full subcategory given by these objects. By definition $\ske$ is skeletal and the inclusion functor is full, faithful and essentially surjective.\qedhere 
\end{proof}
\begin{remark}
	The proof of \Cref{prop:ske} relies on the axiom of choice for classes.
\end{remark}
\begin{remark}
	It is possible to proof that every two skeleta of a given category $\X$ are isomorphic (not only equivalent). For the remaining of this paper we assume that a skeleton $\ske$ of $\X$ and a functor $\pi\colon \X\to \ske$ are chosen once and for all.
\end{remark}

@@

\begin{definition}Let $(\X, \R)$ be a left-linear DPO-rewriting system,  an \emph{abstraction equivalence} between two derivations $\der{D}$ and $\der{D'}$ with the same length and such that $r(\der{D})=r(\der{D}')$, is a family of isomorphisms $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ such that, for every $i\in [0, \lgh(\der{D})-1]$ the following diagram commutes
	\[\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i} \ar@{>->}[l]_{f'_i}&G'_{i+1}\\  L_i \ar[u]^{m'_i} \ar[d]_{m_i}& K_i \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar@{>->}[l]_{l_i} &R_i\ar[u]^{h'_i} \ar[d]_{h_i}\\G_i \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar@{>->}[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{H_i}}}\]

@@

\begin{remark}\label{rem:id}
	Let us consider  $(\{G\}, \alpha, \alpha)$ be a decoration of length $0$ with the same isomorphism as initial and final decoration. Then for every other derivations $(\der{D}, \beta, \omega)$ with source isomorphic to $G$, we have
	\[(\{G\}, \alpha, \alpha)\cdot (\der{D}, \beta, \omega) = (\der{D}, \beta, \omega)\]

@@

	Hence we have
\[
(\{\pi(G_{2,0})\}, \alpha^{-1}_1\circ \omega_1,  (\alpha'_1)^{-1}\circ \omega'_1 )\cdot(\der{D}'_2, \alpha'_2, \omega'_2)  =(\der{D}_2, \alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1\circ \alpha_1^{-1}\circ \omega_1, \omega'_2)\]
Now, by consistency of $\{\phi_{X}\}_{X\in \Deltamin(\der{D}_2)}$ we know that
\[\phi_{G_{2,0}}\circ \alpha_2\circ \omega^{-1}_1\circ \alpha_1=\alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1\]
	therefore we also have the equality
	\[\phi_{G_{2,0}}\circ \alpha_2=\alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1 \circ \alpha^{-1}_1\circ \omega_1 \]
	We conclude that $\{\phi_X\}_{X\in \der{D}_2}$ wintesses the thesis.

@@

\begin{proof} Take two abstraction equivalences $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ and $\{\varphi_X\}_{X\in \Deltamin(\der{D})}$ between $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ and between $(\der{E}, \beta, \xi)$ and $(\der{E}', \beta', \xi')$, respectively. To fix the notation, suppose that goes from $G'$ to $H'$ and $(\der{E}', \beta', \xi')$ from $E'$ to $K'$. We have three cases.

@@

	Vice versa, if $(\{G\}, \alpha, \omega)$ be a derivation of length $0$, then 
	\begin{align*}
	(\{G\}, \alpha, \omega) \cdot (\{G\}, \omega, \alpha)&=(\{G\}, \omega\circ \omega^{-1} \circ \alpha, \alpha )\\&=(\{G\}, \alpha, \alpha)\\
	(\{G\}, \omega, \alpha) \cdot (\{G\}, \alpha, \omega)&=(\{G\}, \alpha\circ \alpha^{-1} \circ \omega, \omega )\\&=(\{G\}, \omega, \omega)
	\end{align*}
	showing that  $[\{G\}, \alpha, \omega]_a$ is an isomorphism.
\end{proof}

@@

\begin{lemma}\label{lem:colim}
	Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, \R)$ a left-linear DPO-rewriting system over it. The following properties hold true:
	\begin{enumerate}
		\item  for eevery derivation $\der{D}$ is a derivation from $G$ to $H$, then the diagram $\Delta(\der{D})$ has a colimit $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ such that $\iota_H$ belongs to $\mathcal{M}$.
		\item Let $\der{D}$ be the concatenation $\der{D}_1\cdot \der{D}_2$ of two derivations $\der{D}_1=\{\dder{D}_{1,i}\}_{i=0}^{n_1}$ between $G$ and $H$ and $\der{D}_2=\{\dder{D}_{2,j}\}_{j=0}^{n_2}$ between $H$ and $T$,  then the colimiting cocone $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ exists too and there is a pushout square
		\[\xymatrix{H\ar[r]^-{\iota_{2, H}} \ar@{>->}[d]_-{\iota_{1, H}} & \tproi{D}{2} \ar[d]^{p_2}\\  \tproi{D}{1} \ar[r]_{p_1}& \tpro{D}}\]
		where $(\tproi{D}{1}, \{\iota_{1, X}\}_{X\in \Delta(\der{D}_1)})$ and $(\tproi{D}{2}, \{\iota_{2, X}\}_{X\in \Delta(\der{D}_2)})$ are the colimiting cocone for $\Delta(\der{D}_1)$ and $\Delta(\der{D}_2)$, respectively.
	\end{enumerate}
\end{lemma}
\begin{remark}\label{rem:cof}
Let $I:\Deltamin(\dder{D})\to \Delta(\dder{D})$ be the inclusion functor. It is immediate to see that such functor is \emph{final} \cite{mac2013categories}. This means that for every functor $F\colon \Delta(\dder{D})\to \Y$ we have:
\begin{enumerate}
	\item if  $(C, \{c_X\}_{X\in \Deltamin(\dder{D})})$ is colimiting for $F\circ I$, then there exists a colimiting cocone $(D, \{d_X\}_{X\in \Delta(\dder{D})})$ for $F$;
	\item $(C, \{c_X\}_{X\in \Deltamin(\dder{D})})$ and $(D, \{d_X\}_{X\in \Delta(\dder{D})})$ are colimiting for, repsectively, $F\circ I$ and $F$, then the canonical arrow $\phi\colon C\to D$ induced by $(D, \{d_X\}_{X\in \Deltamin(\dder{D})})$ is an isomorphism.
\end{enumerate}
\end{remark}

@@

			Let $(C, \{c_X\}_{X\in \Delta(\der{D})})$ be a cocone, since $\Delta(\der{D}')$ is a subdiagram of $\Delta(\der{D})$, we get another cocone $(C, \{c_X\}_{X\in \Delta(\der{D}')})$ which induces an arrow $c'\colon \lpro \der{D}' \rpro \to C$ such that
			\begin{align*}
				c'\circ \iota_{G_n} \circ f_n &=c_{G_n} \circ f_n\\&= c_{D_n}\\&= c_{H}\circ g_n
			\end{align*}
			Therefore the arrows $c'$ and $c_H$ induce a morphism $c\colon P\to C$ and the thesis now follows at once.

@@

\begin{remark}\label{rem:inversa}
	Let  $(\der{D},\alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ be two decorated derivations and $\sigma\colon [0,n]\to [0,n]$ be a consistent permutation between them.  Then its inverse $\sigma^{-1}$ is a consistent permutation between $(\der{D}', \alpha', \omega')$ and $(\der{D},\alpha, \omega)$. Indeed, it is enough to consider, as mediating isomorphism, the inverse $\xi^{-1}_\sigma$ of $\xi_\sigma$.
\end{remark}

@@

\begin{proposition}\label{rem:abscons}
	Let $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ be two decorated derivations in an left-linear rewriting system $(\X, \R)$ of length $l$. Let also $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ be an abstraction equivalence between $\der{D}$ and $\der{D}'$. If $\alpha'\circ \phi_{G_0}=\alpha$ then the the identity $\id{[0,l-1]}$ is a semi-consistent permutation. If, moreover, $\omega'\circ \phi_{G_{l}}=\omega$, then $\id{[0,l-1]}$ is a consistent one.
\end{proposition}
\begin{proof}
	For every $X\in \Deltamin(\der{D}) $, let $X'$ the codomain of $\phi_X$. Then we can define $\xi_{\id{[0,l-1]}}\colon \tpro{D}\to \lpro \der{D}'\rpro$ as the unique morphism such that, for every $X\in \Deltamin(\der{D})$, 
	\[\xi_{\id{[0,l-1]}} \circ \iota_{X}= \iota'_{X'}\circ \phi_X\]

@@

	Now, denoting by $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$, $(\lpro \phi *\der{D} \rpro, \{\iota'_X\}_{X\in \Delta(\phi *\der{D})})$ and $(\lpro \der{D}*\psi \rpro, \{\hat{\iota}_X\}_{X\in \Delta(\der{D}*\psi)})$ the colimiting cocones, in this case we have that the, unique, mediating isomorphisms witnessing the consistency of $\id{[0, \lgh(\der{D})-1]}$ are given by the arrows $\Gamma^{\phi} \colon \tpro{D}\to \lpro \phi*\der{D} \rpro$ and $\Gamma_\psi\colon \tpro{D}\to \lpro \der{D}*\psi \rpro$  such that, for every $X\in \Deltamin(\der{D})$:
	\[\Gamma^\phi \circ \iota_X:=\begin{cases}
		\iota'_{G'}\circ \phi^{-1}  & X=G\\
		\iota'_X & \text{otherwise}
	\end{cases} \qquad \Gamma_\psi \circ \iota_X:=\begin{cases}
		\hat{\iota}_{H'}\circ \psi  & X=H\\
		\hat{\iota}_X & \text{otherwise}
	\end{cases}\]
\end{remark}

>>>>

	A \emph{left-linear DPO-rewriting system} is a pair $(\X, \R)$ where $\X$ is a $\mathcal{M}$-adhesive category and $R$ is a set of left $\mathcal{M}$-linear rules. $(\X, \R)$ is called \emph{linear} if every rule in $R$ is $\mathcal{M}$-linear. Similarly, a \emph{consuming} left-linear DPO-rewriting system is one in which every rule is consuming.

	Given  two objects $G$ and $H$ and a rule $\rho=(l,r)$ in $\R$, a \emph{direct derivation $\mathscr{D}$ from $G$ to $H$ applying the rule $\rho$} is a diagram as the one below, in which both squares are pushouts
	(adhesivity ensures that $f$ belongs to $\mathcal{M}$)

	\[\xymatrix{L \ar[d]_{n}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H}\]
	The arrow $n$ is called the \emph{match} of the derivation, while $h$ is its \emph{back-match}.
	We will denote a direct derivation $\dder{D}$ between $G$ and $H$ as $\dder{D}\colon G\Mapsto H$. 
\end{definition}

@@

\begin{definition}
	Let $(\X, \R)$ be a left-linear DPO-rewriting system, with $\X$ $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of $(\X, \R)$ is the (large)  directed graph $\gpo$ having as vertices the objects of $\X$ and in which an edge between $G$ and $H$ is a direct derivation $\dder{D}\colon G\Mapsto H$.	A \emph{derivation} $\der{D}$ between two objects $G$ and $H$ is a path between them in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are, respectively, $G$ and $H$.
\end{definition}

@@

	Moreover,  if every $\dder{D}_i$ applies the rule $\rho_i\in R$, then we can define an associated sequence of rules $r(\der{D})$ as $\{\rho_i\}_{i=0}^n$.
\end{remark}

@@

\begin{example}\label{ex:1}
	Let us consider the category of $\cat{Graph}$ of graphs and graph morphisms, which is adhesive (see, for instance \cite{lack2005adhesive,lack2006toposes}).   We can endow $\cat{Graph}$ with a set of left-linear rules, for instance taking the following ones.  Here numbers are used to represent the
	morphisms from the interface to the left- and right-hand
	sides. Rules $\rho_0$ and $\rho_1$ are linear: $\rho_0$ generates a
	new node and a new edge, while $\rho_1$ creates a self-loop edge. Instead,
	$\rho_2$ is left-linear but not linear, as it ``merges'' two nodes.
	\begin{center}
		\begin{tikzpicture}[node distance=2mm, font=\small]
			%, baseline=(current bounding box.center), font=\small]
			\node (l) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {};
					\node at (0,0) [node, label=below:$1$] (1) {} ;        
					%
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (l.north) {$L_0$};
			\node [right=of l] (k) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {};
					%
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (k.north) {$K_0$};
			\node[below] at (k.south) {$\rho_0$};
			\node  [right=of k] (r) {
				\begin{tikzpicture}
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {};
					\node at (.5,0) [node, label=below:$2$] (2) {};
					\draw[->] (1) to[out=20, in=160] (2);
					%
					\pgfBox
				\end{tikzpicture}
			};
			\node[font=\scriptsize, above] at (r.north) {$R_0$};
			\path (k) edge[->] node[trans, above] {} (l);
			\path (k) edge[->] node[trans, above] {} (r);    
		\end{tikzpicture}
		%  }
	%
	\hspace{1cm}
	%  \hfill
	%
	%  \subcaptionbox*{}{
		% RULE P1
		%
		\begin{tikzpicture}[node distance=2mm, font=\small]
			\node (l) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {} ;
					% 
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (l.north) {$L_1$};
			\node [right=of l] (k) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {};
					% 
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (k.north) {$K_1$};
			\node[below] at (k.south) {$\rho_1$};
			\node  [right=of k] (r) {
				\begin{tikzpicture}
					\node at (0,0) [node, label=below:$1$] (1) {}
					edge [in=55, out=85, loop] ();
					%
					\pgfBox
				\end{tikzpicture}
			};
			\node[font=\scriptsize, above] at (r.north) {$R_1$};
			\path (k) edge[->] node[trans, above] {} (l);
			\path (k) edge[->] node[trans, above] {} (r);
		\end{tikzpicture}
		%  }
	% 
	\hspace{1cm}
	%\hfill
	%
	%    \subcaptionbox*{}{
		% RULE P2
		%
		\begin{tikzpicture}[node distance=2mm, font=\small]
			\node (l) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {} ;
					\node at (0.5,0) [node, label=below:$2$] (2) {} ;
					%
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (l.north) {$L_2$};
			\node [right=of l] (k) {
				\begin{tikzpicture}
					%
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$1$] (1) {} ;
					\node at (0.5,0) [node, label=below:$2$] (2) {} ;
					%
					\pgfBox
				\end{tikzpicture} 
			};
			\node[font=\scriptsize, above] at (k.north) {$K_2$};
			\node[below] at (k.south) {$\rho_2$};
			\node  [right=of k] (r) {
				\begin{tikzpicture}
					\node at (0,0.53) {}; 
					\node at (0,0) [node, label=below:$12$] (12) {};
					%
					\pgfBox
				\end{tikzpicture}
			};
			\node[font=\scriptsize, above] at (r.north) {$R_2$};
			\path (k) edge[->] node[trans, above] {} (l);
			\path (k) edge[->] node[trans, above] {} (r);
		\end{tikzpicture}
	\end{center}

@@

\begin{remark}
	By definition the inclusion $\ske \to \X$ is an equivalence. In particular, for every objects $X$ of $\X$ there exists $\pi(X)$ in $\ske$ and an isomorphism $\phi_X\colon \pi(X) \to X$.
\end{remark}

\begin{proposition}\label{prop:ske}
	Every category $\X$ has a skeleton. 
\end{proposition}
\begin{proof}
	For every object $X\in \X$, pick a single representative $\pi(X)$ of its isomorphism class. Let $\ske$ be the full subcategory given by these objects. By definition $\ske$ is skeletal and the inclusion functor is full, faithful and essentially surjective.\qedhere 
\end{proof}
\begin{remark}
	The proof of \Cref{prop:ske} relies on the axiom of choice for classes.
\end{remark}
\begin{remark}
	It is possible to prove that every two skeleta of a given category $\X$ are isomorphic (not only equivalent). For the remaining of this paper we assume that a skeleton $\ske$ of $\X$ and a functor $\pi\colon \X\to \ske$ are chosen once and for all.
\end{remark}

@@

\begin{definition}Let $(\X, \R)$ be a left-linear DPO-rewriting system. An \emph{abstraction equivalence} between two derivations $\der{D}$ and $\der{D'}$ with the same length and such that $r(\der{D})=r(\der{D}')$ is a family of isomorphisms $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ such that, for every $i\in [0, \lgh(\der{D})-1]$ the following diagram commutes
	\[\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i} \ar@{>->}[l]_{f'_i}&G'_{i+1}\\  L_i \ar[u]^{m'_i} \ar[d]_{m_i}& K_i \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar@{>->}[l]_{l_i} &R_i\ar[u]^{h'_i} \ar[d]_{h_i}\\G_i \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar@{>->}[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{H_i}}}\]

@@

\begin{remark}\label{rem:id}
	Let  $(\{G\}, \alpha, \alpha)$ be a decoration of length $0$ with the same isomorphism as initial and final decoration. Then for every other derivations $(\der{D}, \beta, \omega)$ with source isomorphic to $G$, we have
	\[(\{G\}, \alpha, \alpha)\cdot (\der{D}, \beta, \omega) = (\der{D}, \beta, \omega)\]

@@

	Hence we have
\[
(\{\pi(G_{2,0})\}, \alpha^{-1}_1\circ \omega_1,  (\alpha'_1)^{-1}\circ \omega'_1 )\cdot(\der{D}'_2, \alpha'_2, \omega'_2)  =(\der{D}_2, \alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1\circ \alpha_1^{-1}\circ \omega_1, \omega'_2)\]
Now, by consistency of $\{\phi_{X}\}_{X\in \Deltamin(\der{D}_2)}$ we know that
\[\phi_{G_{2,0}}\circ \alpha_2\circ \omega^{-1}_1\circ \alpha_1=\alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1\]
	therefore we also have the equality
	\[\phi_{G_{2,0}}\circ \alpha_2=\alpha'_2\circ (\omega'_1)^{-1} \circ \alpha'_1 \circ \alpha^{-1}_1\circ \omega_1 \]
	We conclude that $\{\phi_X\}_{X\in \der{D}_2}$ witnesses the thesis.

@@

\begin{proof} Take two abstraction equivalences $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ and $\{\varphi_X\}_{X\in \Deltamin(\der{D})}$ between $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ and between $(\der{E}, \beta, \xi)$ and $(\der{E}', \beta', \xi')$, respectively. To fix the notation, suppose that $(\der{D}', \alpha', \omega')$ goes from $G'$ to $H'$ and $(\der{E}', \beta', \xi')$ from $E'$ to $K'$. We have three cases.

@@

	Vice versa, if $(\{G\}, \alpha, \omega)$ is a derivation of length $0$, then 
	\begin{align*}
	(\{G\}, \alpha, \omega) \cdot (\{G\}, \omega, \alpha)&=(\{G\}, \omega\circ \omega^{-1} \circ \alpha, \alpha )\\&=(\{G\}, \alpha, \alpha)\\
	(\{G\}, \omega, \alpha) \cdot (\{G\}, \alpha, \omega)&=(\{G\}, \alpha\circ \alpha^{-1} \circ \omega, \omega )\\&=(\{G\}, \omega, \omega)
	\end{align*}
	showing that  $[\{G\}, \alpha, \omega]_a$ is an isomorphism.
\end{proof}

@@

\begin{lemma}\label{lem:colim}
	Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, \R)$ a left-linear DPO-rewriting system over it. The following properties hold true:
	\begin{enumerate}
		\item  if $\der{D}$ is a derivation from $G$ to $H$, then the diagram $\Delta(\der{D})$ has a colimit $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ such that $\iota_H$ belongs to $\mathcal{M}$.
		\item if $\der{D}$ is the concatenation $\der{D}_1\cdot \der{D}_2$ of two derivations $\der{D}_1=\{\dder{D}_{1,i}\}_{i=0}^{n_1}$ between $G$ and $H$ and $\der{D}_2=\{\dder{D}_{2,j}\}_{j=0}^{n_2}$ between $H$ and $T$,  then the colimiting cocone $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ exists too and there is a pushout square
		\[\xymatrix{H\ar[r]^-{\iota_{2, H}} \ar@{>->}[d]_-{\iota_{1, H}} & \tproi{D}{2} \ar[d]^{p_2}\\  \tproi{D}{1} \ar[r]_{p_1}& \tpro{D}}\]
		where $(\tproi{D}{1}, \{\iota_{1, X}\}_{X\in \Delta(\der{D}_1)})$ and $(\tproi{D}{2}, \{\iota_{2, X}\}_{X\in \Delta(\der{D}_2)})$ are the colimiting cocone for $\Delta(\der{D}_1)$ and $\Delta(\der{D}_2)$, respectively.
	\end{enumerate}
\end{lemma}
\begin{remark}\label{rem:cof}
Let $I:\Deltamin(\dder{D})\to \Delta(\dder{D})$ be the inclusion functor. It is immediate to see that such functor is \emph{final} \cite{mac2013categories}. This means that for every functor $F\colon \Delta(\dder{D})\to \Y$ we have:
\begin{enumerate}
	\item if  $(C, \{c_X\}_{X\in \Deltamin(\dder{D})})$ is colimiting for $F\circ I$, then there exists a colimiting cocone $(D, \{d_X\}_{X\in \Delta(\dder{D})})$ for $F$;
	\item $(C, \{c_X\}_{X\in \Deltamin(\dder{D})})$ and $(D, \{d_X\}_{X\in \Delta(\dder{D})})$ are colimiting for, respectively, $F\circ I$ and $F$, then the canonical arrow $\phi\colon C\to D$ induced by $(D, \{d_X\}_{X\in \Deltamin(\dder{D})})$ is an isomorphism.
\end{enumerate}
\end{remark}

@@

			Let $(C, \{c_X\}_{X\in \Delta(\der{D})})$ be a cocone. Since $\Delta(\der{D}')$ is a subdiagram of $\Delta(\der{D})$, we get another cocone $(C, \{c_X\}_{X\in \Delta(\der{D}')})$ which induces an arrow $c'\colon \lpro \der{D}' \rpro \to C$ such that
			\begin{align*}
				c'\circ \iota_{G_n} \circ f_n &=c_{G_n} \circ f_n\\&= c_{D_n}\\&= c_{H}\circ g_n
			\end{align*}
			Therefore the arrows $c'$ and $c_H$ induce a morphism $c\colon P\to C$ and the thesis now follows at once.

@@

\begin{remark}\label{rem:inversa}
	Let  $(\der{D},\alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ be two decorated derivations and $\sigma\colon [0,n]\to [0,n]$ a consistent permutation between them.  Then its inverse $\sigma^{-1}$ is a consistent permutation between $(\der{D}', \alpha', \omega')$ and $(\der{D},\alpha, \omega)$. Indeed, it is enough to consider, as mediating isomorphism, the inverse $\xi^{-1}_\sigma$ of $\xi_\sigma$.
\end{remark}

@@

\begin{proposition}\label{rem:abscons}
	Let $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ be two decorated derivations in a left-linear rewriting system $(\X, \R)$ of length $l$. Let also $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ be an abstraction equivalence between $\der{D}$ and $\der{D}'$. If $\alpha'\circ \phi_{G_0}=\alpha$ then the the identity $\id{[0,l-1]}$ is a semi-consistent permutation. If, moreover, $\omega'\circ \phi_{G_{l}}=\omega$, then $\id{[0,l-1]}$ is a consistent one.
\end{proposition}
\begin{proof}
	For every $X\in \Deltamin(\der{D}) $, let $X'$ the codomain of $\phi_X$. Then we can define $\xi_{\id{[0,l-1]}}\colon \tpro{D}\to \lpro \der{D}'\rpro$ as the unique morphism such that, for every $X\in \Deltamin(\der{D})$, 
	\[\xi_{\id{[0,l-1]}} \circ \iota_{X}= \iota'_{X'}\circ \phi_X\]

@@

	Now, denoting by $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$, $(\lpro \phi *\der{D} \rpro, \{\iota'_X\}_{X\in \Delta(\phi *\der{D})})$ and $(\lpro \der{D}*\psi \rpro, \{\hat{\iota}_X\}_{X\in \Delta(\der{D}*\psi)})$ the colimiting cocones, in this case we have that the (unique) mediating isomorphisms witnessing the consistency of $\id{[0, \lgh(\der{D})-1]}$ are given by the arrows $\Gamma^{\phi} \colon \tpro{D}\to \lpro \phi*\der{D} \rpro$ and $\Gamma_\psi\colon \tpro{D}\to \lpro \der{D}*\psi \rpro$  such that, for every $X\in \Deltamin(\der{D})$:
	\[\Gamma^\phi \circ \iota_X:=\begin{cases}
		\iota'_{G'}\circ \phi^{-1}  & X=G\\
		\iota'_X & \text{otherwise}
	\end{cases} \qquad \Gamma_\psi \circ \iota_X:=\begin{cases}
		\hat{\iota}_{H'}\circ \psi  & X=H\\
		\hat{\iota}_X & \text{otherwise}
	\end{cases}\]
\end{remark}
